Nearest Neighbour Methods for Length of Stay Prediction

### 2015-02-14 papers/paper.tex
Small changes to headings

$k$-NN allows all features to be taken
into account equally when classifying a new example. However, since feature
values often span different ranges and units of measurement, a particularly
large range of values for a feature would contribute much more to the distance
than another feature that spans a lower set of values.
To avoid this, it is important to \textit{normalise} the
values of each numeric feature before using $k$-NN: for each feature $i$, we
calculate the normalised value of the feature for each training example. This
is done using the following relationship:
\begin{equation}
\label{eq:normalise}
a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
\end{equation}
where $a_i$ is the normalised feature value, $v_i$ is the actual value in the
data set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is
a scaling factor and $t$ indicates how much to translate the range by. If $s=1$
and $t=0$, then the range of normalised values will be in $[0,1]$.

@@

\subsubsection{Ranked Distance Nearest Neighbour}
\paragraph{Intuition}
As we mentioned earlier, all features are considered equally in a $k$-NN
classifier, which implies that all features are equally important in predicting
the class. In practice, however, this is not always the case: to predict the
LOS of a trauma patient, the severity of their injury should affect the final
LOS more than whether or not they can speak English (which could be just two
of many features that are recorded about a patient). Our contribution is a
new NN algorithm that uses a modified distance function which takes into
account the relative importance of all features called Ranked Distance
Nearest Neighbour (RD).

\paragraph{Mathematical formulation}
Given two instances $\mathbf{x} = (x_1,x_2,\ldots,x_n)$ and
$\mathbf{y} = (y_1,y_2,\ldots,y_n)$, the distance between them is given by:
\begin{equation*}
\mathrm{Distance} = \sum_{i=1}^n w_i |x_i-y_i|
\end{equation*}
where the $w_i$s weight the contribution of the $i$-th feature to the overall
distance that is computed by the $k$-NN algorithm. We assume that all feature
values have been normalised, as described above.

\paragraph{Assignment of weights}
The weights $w_i$ can be tuned to match the particular problem that is being
investigated. For our LOS classification problem, we will consider one way
of weighting features.
First, we compute the correlation coefficients as described in the point above.
This gives us a ranking of the importance of each feature to the prediction of
the class, with 1 being the highest rank (indicating the most importance) and
$n$ being the lowest rank (indicating the least importance).
Instead of using the values of the correlation directly to weight
the distance contribution of each feature, we specify a function:
\begin{equation*}
f : \mathrm{Rank} \rightarrow \mathbb{R}, \mathrm{Rank} \in \{1,2,\ldots,n\}
\end{equation*}
that describes
how the weights of the features vary with the rank. This function should
inituitively be non-increasing and should produce lower values when the rank
number is greater, meaning that less important features have a lower weight
in the distance calculation.
Note that choosing $f(\mathrm{Rank}) = k$ for any non-zero constant $k$
results in all features contributing equally to the distance calculation.
We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}^c}$, where
$c$ varies from 0 to 1.

>>>>

\subsection{Normalisation}
$k$-NN allows all features to be taken
into account equally when classifying a new example. However, since feature
values often span different ranges and units of measurement, a particularly
large range of values for a feature would contribute much more to the distance
than another feature that spans a lower set of values.
To avoid this, it is important to \textit{normalise} the
values of each numeric feature before using $k$-NN: for each feature $i$, we
calculate the normalised value of the feature for each training example. This
is done using the following relationship:
\begin{equation}
\label{eq:normalise}
a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
\end{equation}
where $a_i$ is the normalised feature value, $v_i$ is the actual value in the
data set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is
a scaling factor and $t$ indicates how much to translate the range by. If $s=1$
and $t=0$, then the range of normalised values will be in $[0,1]$.

@@

\subsubsection{Ranked Distance Nearest Neighbour}
\paragraph{Intuition.}
As we mentioned earlier, all features are considered equally in a $k$-NN
classifier, which implies that all features are equally important in predicting
the class. In practice, however, this is not always the case: to predict the
LOS of a trauma patient, the severity of their injury should affect the final
LOS more than whether or not they can speak English (which could be just two
of many features that are recorded about a patient). Our contribution is a
new NN algorithm that uses a modified distance function which takes into
account the relative importance of all features called Ranked Distance
Nearest Neighbour (RD).

\paragraph{Mathematical formulation.}
Given two instances $\mathbf{x} = (x_1,x_2,\ldots,x_n)$ and
$\mathbf{y} = (y_1,y_2,\ldots,y_n)$, the distance between them is given by:
\begin{equation*}
\mathrm{Distance} = \sum_{i=1}^n w_i |x_i-y_i|
\end{equation*}
where the $w_i$s weight the contribution of the $i$-th feature to the overall
distance that is computed by the $k$-NN algorithm. We assume that all feature
values have been normalised, as described above.

\paragraph{Assignment of weights.}
The weights $w_i$ can be tuned to match the particular problem that is being
investigated. For our LOS classification problem, we will consider one way
of weighting features.
First, we compute the correlation coefficients as described in the point above.
This gives us a ranking of the importance of each feature to the prediction of
the class, with 1 being the highest rank (indicating the most importance) and
$n$ being the lowest rank (indicating the least importance).
Instead of using the values of the correlation directly to weight
the distance contribution of each feature, we specify a function:
\begin{equation*}
f : \mathrm{Rank} \rightarrow \mathbb{R}, \mathrm{Rank} \in \{1,2,\ldots,n\}
\end{equation*}
that describes
how the weights of the features vary with the rank. This function should
inituitively be non-increasing and should produce lower values when the rank
number is greater, meaning that less important features have a lower weight
in the distance calculation.
Note that choosing $f(\mathrm{Rank}) = k$ for any non-zero constant $k$
results in all features contributing equally to the distance calculation.
We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}^c}$, where
$c$ varies from 0 to 1.
